### 2018-08-01 conclusion.tex
Backup

The pipelines presented in chapter \ref{chap:pipeline} and the MaxQuant suite offered the easiest to deploy and most output rich of all the cost-free pieces of software available for the proteomics community. All of them have pros and cons, and their combined usage was found to provide the best results, underlining the importance of the development of programs implementing open data formats exchangeable across tools. More functionalities can be added, including \textit{de novo} and \ac{PTM} search, and missing data management. BayesQuant, the probabilistic framework for relative quantification, presented in chapter \ref{chap:model} can be passed any peptide level dataset and provide estimates of the uncertainty behind traditional \ac{log2FC} estimates and nuisance effects. It can be customised to expand its currently implemented model. The lack of success in the attempt to model the peptide bias using the aminoacidic sequence reveals the complexity of the competitive ionization problem. Recently developed tools like RawQuant could be applied in future work to extract from RAW files a feature-rich representation of the collected spectra that could be used to model this phenomenon with a Deep Learning approach. The results from chapter \ref{chap:benchmark} manifest that \ac{NZ} can at the same time decrease computational costs and access the state of the art in proteomics research by making use of the latest open-source software, and it is encouraged to continue on that line.

>>>>

The pipelines presented in chapter \ref{chap:pipeline} offer an output-rich and easy to deploy cost-free pieces of software. All of them have pros and cons, and their combined usage was found to provide the best results, underlining the importance of the development of programs implementing open data formats exchangeable across tools. More functionalities can be added, including \textit{de novo} and \ac{PTM} search, and missing data management. BayesQuant, the probabilistic framework for relative quantification, presented in chapter \ref{chap:model} can be passed any peptide level dataset and provide estimates of the uncertainty behind frequentist \ac{log2FC} estimates and nuisance effects. It can be customised to expand its currently implemented model. The lack of success in the attempt to model the peptide bias using the aminoacidic sequence reveals the complexity of its origin. Recently developed tools like RawQuant could be applied in future work to extract a feature-rich representation of the collected spectra that could be used to model this phenomenon with a Deep Learning approach. The results from chapter \ref{chap:benchmark} manifest that \ac{NZ} can at the same time decrease computational costs and access the state of the art in proteomics research by making use of the latest open-source software, and it is encouraged to continue on that line.

### 2018-08-01 modelling.tex
Backup

A mass spectrometry-proteomics workflow aims at providing estimates of the treatment effect, upon which the biological interpretation of the results is done, while minimising the remaining unwanted effects. A mathematical framework implemented to model these effects, similar to the one used by MSqRob:

@@

\begin{figure}[!h]
\centering
\begin{subfigure}{.9\textwidth}
\centering
\caption*{Pooled}
\includegraphics[width=0.85\linewidth]{model_pooled}
\end{subfigure}
\bigskip
\begin{subfigure}{.9\textwidth}
\centering
\caption*{Unpooled}
\includegraphics[width=0.85\linewidth]{model_unpooled}
\end{subfigure}
\bigskip
\begin{subfigure}{.9\textwidth}
\centering
\caption*{Hierarchical}
\includegraphics[width=0.85\linewidth]{model_hierarchical}
\end{subfigure}
\caption[]{Alternative multilevel modelling approaches. A pooled model assumes that the parameter distribution is the same for all the datapoints.  The unpooled model represents the opposite case, every datapoint is modelled by a different probability distribution. Finally, a hierarchical model settles for a middle ground where the parameters will not be exactly the same but not completely different either. This is achieved by generating a global distribution from which the parameter for each datapoint is sampled \footnotemark{}.}
\label{fig:multilevel}
\end{figure}

\footnotetext{\href{https://docs.pymc.io/notebooks/multilevel\_modeling.html}{https://docs.pymc.io/notebooks/multilevel\_modeling.html}}

This way, independent and identically distributed (\ac{IID}) priors were set for the intercept, treatment, peptide and run effects. The particular effect observed on each peptide was then modelled as a value sampled from the corresponding \ac{IID} prior. A diagram of the resulting model is displayed in figure \ref{fig:daft_model}).

@@

The \ac{VI} mean-field strategy was applied to approximate the posterior probability distribution of the \ac{log2FC}. \ac{ELBO} optimisation was run for 40k iterations. Optimisation was validated by visual inspection of \ac{ELBO} traceplots over the 40k iterations. Once optimisation was validated, 10k samples were drawn from the approximate distribution to simulate the true posterior.

@@

The code below formalizes in PyMC3 code the bias in \ac{MS1} intensity measurements due to a random effect.

@@

The snippet above loads the program, reads in the data, selects the data for a specific protein, and builds a model that matches the number of peptides observed. The posterior distribution can be computed using pure MCMC or VI methods, as stated in section \ref{subsec:posterior_compute}. The VI procedure will now be explained, due to its significantly better performance.

@@

A step by step exposition of the plots and results produced by the program when ran on the proteins  P0A818 (\textit{E. coli}) and A6NDG6 (\textit{Homo sapiens}) will follow.

@@

After the \texttt{.fit()} method is run, a plot of the maximisation of the \ac{ELBO} is produced automatically. Even though the ELBO took extremely low (negative) values at start (see figure \ref{fig:ELBO}), it was maximised fast to values around 0, indicating that the mean-field approximation was acceptably good.

@@

The study of the \ac{log2FC} posterior probability distribution (see figure \ref{fig:posteriors}) shows the most likely parameter values and empowers Bayesian Null Hypothesis Testing. The 95\% High Probability Density Interval (\ac{HPDI}), which is the narrowest interval containing 95\% of the total probability, indicates the most likely true values of the parameter. Presence or absence of overlap between the 95\% \ac{HPDI} and the Region Of Practical Equivalence (\ac{ROPE}) provides a formal way of discarding a point value of the parameter. The \ac{ROPE} is a small interval considered to be essentually the same as the null value \cite{Kruschke}. In this and all future applications in the thesis, the null value is set to 0, and the interval is given a width of 0.4 on each side.

\begin{figure}[!h]
\centering
\begin{subfigure}{0.8\textwidth}
\caption*{A}
\includegraphics[width=\linewidth]{posteriors/P0A8I8}
\end{subfigure}
\bigskip
\begin{subfigure}{0.8\textwidth}
\caption*{B}
\includegraphics[width=\linewidth]{posteriors/A6NDG6}
\end{subfigure}
\caption[Posteriors for 2 proteins]{Annotated posterior probability distribution for the \ac{log2FC} estimate for proteins P0A818 (\textit{E. coli}, \textbf{A}) and A6NDG6 (\textit{Homo sapiens}, \textbf{B}). The bar height is mapped to the probability mass in the corresponding interval. The 95\% \ac{HPDI} is marked with a black line. The \ac{ROPE} defined as a 0.4 window around 0 is shown in red. A green vertical line marks the expected \ac{log2FC} estimate for \textit{E. coli} proteins ($\log_2(3)$).}
\label{fig:posteriors}
\end{figure}

The analysis indicated that P0Q8I8 (\textit{E. coli}) most likely has a \ac{log2FC} between 1.4 and 1.7, and it\textquotesingle significantly different from 0, as no overlap was observed between \ac{HPDI} and \ac{ROPE}. On the other hand, the human protein was found to acquire a \ac{log2FC} most likely between -0.1 and 0.192. This range is fully contained within the \ac{ROPE}, which indicates that the \ac{log2FC} was practically 0 (see figure \ref{fig:posteriors}).

@@

\begin{figure}[!h]
\centering
\begin{subfigure}{0.8\textwidth}
\caption*{A}
\includegraphics[width=\linewidth]{PPC/eps/PCA_P0A8I8}
\end{subfigure}
\bigskip
\begin{subfigure}{0.8\textwidth}
\caption*{B}
\includegraphics[width=\linewidth]{PPC/eps/PCA_A6NDG6}
\end{subfigure}
\caption[Posterior predictive checks]{Projection of the posterior predictive checks on the 2D plane capturing most variance. Simulated datapoints are shown in blue, whereas actual observations are shown in red.}
\label{fig:ppc}
\end{figure}

@@

Indeed, a general trend of width reduction was observed with the increased amount of peptides. For example, the widest intervals are observed on proteins with just two peptides available, and the narrowest on proteins with ten. Moreover, the \ac{HPDI} tended to align more strongly around an agreement value as well.

@@

where $\theta_k$ is the weight of the $k^{th}$ feature, and $x_k$ is the numerical value of the $k^{th}$ feature. $\sigma^{\theta}$ is the prior for the standard deviation of the Normal all feature weights are sampled from.

@@

While BayesQuant performed well on the majority of proteins, it took around 30 seconds to approximate each protein\textquotesingle posterior distributions using \ac{VI}, and more than two minutes with \ac{MCMC} methods. This translates to a dataset of 1k proteins taking 500 minutes (>8h) to be processed. However, the average modern computer comes with several processors that could support the parallelization of the program in several threads, decreasing the total computing time.

@@

Besides the predefinition of a \ac{ROPE} and the check of overlap with the 95 \% \ac{HPDI}, more advanced model comparison methods making use of the Bayes factor are available. The Bayes factor is defined as the ratio of the posterior probability of the models given the data \cite{Kruschke}. It could be used to measure how much more likely a model taking into account a treatment effect is compared to another one where the treatment effect is considered null i.e the \ac{log2FC} is set to 0. The result would provide an alternative criteria for the declaration of a protein being differentially abundant.

@@

Unfortunately, during the development of the project, it was acknowledged that the peptide effect is a very difficult phenomenon to capture in a simple Bayesian model as the one presented in this work. This is due to the motley and and multilevel nature of the noise caused by the aminoacidic sequences:

@@

\item The mobile proton model \cite{Boyd2010}, which states that the peptide\textquotesingle s fragmentation pattern changes drastically depending on the number of charges and its distribution on the sequence. The charge distribution is hypothesised to be in turn determined by the position of positive residues that can allocate the charges. This phenomenon implies that the \ac{PSM} step is more difficult as a different spectrum is expected for molecules with the same peptide sequence but different charges. The decreased identifications and quality of the peaks makes feature extraction algorithms like the Apex module in moFF would extract worse \ac{MS1} intensities, in turn injecting noise in the quantification data.

\item The characterization of powerful sequence properties that specifically capture the differential ionization efficiency is undone work, as most feature extractors are designed to solve different problems, like protein sructure prediction.

@@

Modelling such a complex phenomenon will require accordingly complex predictive architectures and the introduction of proper neural networks, capable of deciphering the intrincate pattern mapping sequence to noise in the spectrometer\textquotesingle s measurements.

@@

The characterization of the uncertainty behind a \ac{log2FC} estimate can be used to better inform downstream analysis programs and eventually help interpret the biological results. For example, a functional analysis program could decide to drop proteins declared differentially abundant if the uncertainty is greater than a threshold. A Gene Set Enrichment Analysis (\ac{GSEA}) could make use of the probabilistic information to refine the results of the enrichment.

@@

A Bayesian framework for the relative quantification of protein abundance rations (Fold-Change, FC) using \ac{MS1} intensity was presented in this work. Unlike the currently prevailing methods in the field, it provides uncertainty estimates that provide a direct interpretation about the accuracy of the quantification. Moreover, preliminary results on the sequence modelling of the peptide bias revealed that more complex architectures and sequnce features are required to solve this problem. Finally, the method can be used to investigate the presence of batch effects in \ac{MS} experiments for their minimisation in future. While the software can be integrated in any pipeline producing peptide-level data, further testing and improvements are probably required for optimal performance.

>>>>

A mass spectrometry-proteomics workflow aims at providing estimates of the treatment effect, upon which the biological interpretation of the results is done, while minimising the remaining unwanted effects. A mathematical modelling of these effects (see equation \ref{eq:model}), similar to the one used by MSqRob, can be deployed to attempt to minimise the noise:

@@

\begin{figure}[!h]
\centering
\begin{subfigure}{.9\textwidth}
\centering
\caption*{Pooled}
\includegraphics[width=0.85\linewidth]{model_pooled}
\end{subfigure}
\bigskip
\begin{subfigure}{.9\textwidth}
\centering
\caption*{Unpooled}
\includegraphics[width=0.85\linewidth]{model_unpooled}
\end{subfigure}
\bigskip
\begin{subfigure}{.9\textwidth}
\centering
\caption*{Hierarchical}
\includegraphics[width=0.85\linewidth]{model_hierarchical}
\end{subfigure}
\caption[]{Alternative multilevel modelling approaches. A pooled model assumes that the parameter distribution is the same for all the datapoints.  The unpooled model represents the opposite case, every datapoint is modelled by a different probability distribution. Finally, a hierarchical model settles for a middle ground where the parameters will not be exactly the same but not completely different either. This is achieved by generating a global distribution from which the parameter for each datapoint is sampled \footnotemark{}.}
\label{fig:multilevel}
\end{figure}
\footnotetext{\href{https://docs.pymc.io/notebooks/multilevel\_modeling.html}{https://docs.pymc.io/notebooks/multilevel\_modeling.html}}

This way, independent and identically distributed (\ac{IID}) priors were set for the intercept, treatment, peptide and run effects. The particular effect observed on each peptide was then modelled as a value sampled from the corresponding \ac{IID} prior. A diagram of the resulting model is displayed in figure \ref{fig:daft_model}.

@@

The \ac{VI} mean-field strategy was applied to approximate the posterior probability distribution of the \ac{log2FC}. \ac{ELBO} optimisation was run for 40k iterations. Optimisation was validated by visual inspection of \ac{ELBO} traceplots. Once optimisation was validated, 10k samples were drawn from the approximate distribution to simulate the true posterior.

@@

A PyMC3 implementation of the priors set on the bias in \ac{MS1} intensity measurements due to a random effect is shown here.

@@

The snippet above loads the program, reads in the data, selects a specific protein, and builds a model that matches the number of peptides observed. The posterior distribution can be computed using pure MCMC or VI methods, as stated in section \ref{subsec:posterior_compute}. The VI procedure will now be explained, due to its significantly better performance in this particular problem.

@@

A step by step exposition of the plots and results produced by the program when ran on the proteins  P0A818 (\textit{E. coli}) and A6NDG6 (\textit{Homo sapiens}) follows.

@@

A plot of the maximisation of the \ac{ELBO} is automatically produced when the texttt{.fit()} method is run. Even though the ELBO took extremely low (negative) values at start (see figure \ref{fig:ELBO}), it was maximised fast to values around 0, indicating that the mean-field approximation was acceptably good.

@@

The study of the \ac{log2FC} posterior probability distribution (see figure \ref{fig:posteriors}) shows the most likely parameter values and empowers Bayesian Null Hypothesis Testing. The 95\% High Probability Density Interval (\ac{HPDI}), which is the narrowest interval containing 95\% of the total probability, indicates the most likely true values of the parameter. Presence or absence of overlap between the 95\% \ac{HPDI} and the Region Of Practical Equivalence (\ac{ROPE}) provides a formal way of discarding a point value of the parameter. The \ac{ROPE} is a small interval considered to be essentually the same as the null value \cite{Kruschke}. In this and all future applications in the thesis, the null value is set to 0, and the \ac{ROPE} is arbitrality set to $[-0.4, 0.4]$ based on the data.

\begin{figure}[!h]
\centering
\begin{subfigure}{0.8\textwidth}
\caption*{A}
\includegraphics[width=\linewidth]{posteriors/P0A8I8}
\end{subfigure}
\bigskip
\begin{subfigure}{0.8\textwidth}
\caption*{B}
\includegraphics[width=\linewidth]{posteriors/A6NDG6}
\end{subfigure}
\caption[Posteriors for 2 proteins]{Annotated posterior probability distribution for the \ac{log2FC} estimate for proteins P0A818 (\textit{E. coli}, \textbf{A}) and A6NDG6 (\textit{Homo sapiens}, \textbf{B}). The bar height is mapped to the probability mass in the corresponding interval. The 95\% \ac{HPDI} is marked with a black line. The \ac{ROPE} defined as a 0.4 window around 0 is shown in red. A green vertical line marks the expected \ac{log2FC} estimate for \textit{E. coli} proteins ($\log_2(3)=1.58$).}
\label{fig:posteriors}
\end{figure}

The analysis indicated that P0Q8I8 (\textit{E. coli}) most likely has a \ac{log2FC} between 1.4 and 1.7, and was found to be significantly different from 0, as no overlap was observed between the \ac{HPDI} and the\ac{ROPE}. On the other hand, the human protein was found to acquire a \ac{log2FC} most likely between -0.1 and 0.2. This range is fully contained within the \ac{ROPE}, which indicates that the \ac{log2FC} was practically 0 (see figure \ref{fig:posteriors}).

@@

\begin{figure}[!h]
\centering
\begin{subfigure}{0.8\textwidth}
\caption*{A}
\includegraphics[width=\linewidth]{PPC/eps/PCA_P0A8I8}
\end{subfigure}
\bigskip
\begin{subfigure}{0.8\textwidth}
\caption*{B}
\includegraphics[width=\linewidth]{PPC/eps/PCA_A6NDG6}
\end{subfigure}
\caption[Posterior predictive checks]{Projection of the posterior predictive checks on the 2D plane capturing most variance. Simulated datapoints are shown in blue, whereas actual observations are shown on a clear color.}
\label{fig:ppc}
\end{figure}

@@

Indeed, a general trend of width reduction was observed as the amount of peptides increased. For example, the widest intervals are observed on proteins with just two observed peptides, and the narrowest on proteins with ten. Moreover, the \ac{HPDI}s tended to align more strongly around an agreement value as well.

@@

where $\theta_k$ is the weight of the $k^{th}$ feature, and $x_k$ is the numerical value of the $k^{th}$ feature. $\sigma^{\theta}$ is the prior for the standard deviation of the distribution all feature weights are sampled from.

@@

While BayesQuant performed well on the majority of proteins, it took around 30 seconds to approximate each protein\textquotesingle s posterior distributions using \ac{VI}, and more than two minutes with \ac{MCMC} methods. This translates to a dataset of 1k proteins taking 500 minutes (>8h) to be processed. However, the average modern computer comes with several processors that could support the parallelization of the program in several threads, decreasing the total computing time.

@@

Besides the predefinition of a \ac{ROPE} and the check of overlap with the 95 \% \ac{HPDI}, more advanced model comparison methods making use of the Bayes factor are available. The Bayes factor is defined as the ratio of the posterior probability of the models given the data \cite{Kruschke}. It could be used to measure how much more likely a model taking into account a treatment effect is compared to another one where the treatment effect is considered null i.e the \ac{log2FC} is set to 0. The result would provide an alternative criterion for the declaration of a protein being differentially abundant.

@@

Unfortunately, during the development of the project, it was acknowledged that the peptide effect is a very difficult phenomenon to capture in a simple Bayesian model as the one presented in this work. This is due to the motley and multilevel nature of the noise caused by the aminoacidic sequences:

@@

\item The mobile proton model \cite{Boyd2010}, which states that the peptide\textquotesingle s fragmentation pattern changes drastically depending on the number of charges and its distribution on the sequence. The charge distribution is hypothesised to be in turn determined by the position of positive residues that can allocate the charges. This phenomenon implies that the \ac{PSM} step is more difficult as a different spectrum is expected for molecules with the same peptide sequence but different charges. The decreased identifications and quality of the peaks worsen the performance of feature extraction algorithms like the Apex module in moFF. In turn, this results in the injection of more noise in the quantification data.

\item The characterization of powerful sequence properties that specifically capture differential ionization efficiency is undone work, as most feature extractors are designed to solve different problems, like protein sructure prediction.

@@

Modelling such a complex phenomenon will require accordingly complex predictive architectures and the introduction of proper neural networks, capable of deciphering the intricate pattern mapping sequence to noise in the spectrometer\textquotesingle s measurements.

@@

The characterization of the uncertainty behind a \ac{log2FC} estimate can be used to better inform downstream analysis programs and eventually help interpret the biological results. For example, a functional analysis program could decide to drop proteins declared differentially abundant if the uncertainty is greater than a threshold. A Gene Set Enrichment Analysis (\ac{GSEA}) could make use of the probabilistic information to refine its results.

@@

A Bayesian framework for the relative quantification of protein abundance rations (Fold-Change, FC) using \ac{MS1} intensity was presented in this chapter. Unlike the currently prevailing methods in the field, it gives uncertainty estimates that provide a direct interpretation about the accuracy of the quantification. Moreover, preliminary results on the sequence modelling of the peptide bias revealed that more complex architectures and sequnce features are required to model noise with peptide sequences. Finally, the method can be used to investigate the presence of batch effects in \ac{MS} experiments for their minimisation in future experiments. While the software can be integrated in any pipeline producing peptide-level data, further testing and improvements are probably required for optimal performance.

### 2018-08-01 pipeline.tex
Backup

The extracted apex MS1 intensities were used as proxy for peptide abundance to estimate protein $\log_2$(ratios), or \textit{log2-fold-changes} (\ac{log2FC}) across condition H and condition L. During quantification, it is important to take into account as many of the effects influencing protein quantification. In this case, besides the different H and L conditions, peptide, run and fraction effects could be distinguished. The MSqRob quantification engine treats effects differently depending on whether or not they are random, or fixed. Fixed effects should have a consistent impact in the ion currents measured in the spectrometer, whereas random effects are truly random and thus unpredictable.

@@

The low attained matching rate manifests existing room for improvement in the currently available tools. As explained in section \ref{sec:search_engines}, it has been shown that the combined usage of multiple search engines can increase identifications, since the different statistical frameworks implemented in each of them compensate each other\textquotesingle s caveats. Only MS-GF+ was used in thi work for simplicity. Furthermore, \textit{de novo} search engines are available and supported by SearchGUI, and the latest developments in the field, like the IdentiPy engine \cite{Levitsky2018} are progressively incorporated to the tool \footnote{\href{https://github.com/compomics/peptide-shaker/issues/309}{https://github.com/compomics/peptide-shaker/issues/309}}. Their usage is predicted to improve identification rates and quality of the matches.

@@

Sample fractionation allows for a deeper analysis of the peptide mix, which in turn leads to increased identifications and an increased amount of data to be analyzed. However, it psimultaneously provides an extra random effect that difficults quantification if not handled properly.

>>>>

The extracted apex MS1 intensities were used as proxy for peptide abundance to estimate protein $\log_2$(ratios), or logarithm base 2 fold changes, $\log_2(FC)$ (\ac{log2FC}) across conditions H and L. During quantification, it is important to take into account as many of the effects influencing protein quantification. In this case, besides the different H and L conditions, peptide, run and fraction effects could be distinguished. The MSqRob quantification engine treats effects differently depending on whether or not they are random or fixed. Fixed effects should have a consistent impact in the ion currents measured in the spectrometer, whereas random effects are truly random and thus unpredictable.

@@

The low attained matching rate manifests existing room for improvement in the currently available tools. As explained in section \ref{sec:search_engines}, it has been shown that the combined usage of multiple search engines can increase identifications, since the different statistical frameworks implemented in each of them compensate each other\textquotesingle s caveats. Only MS-GF+ was used in this work for simplicity. Furthermore, \textit{de novo} search engines are available and supported by SearchGUI, and the latest developments in the field, like the IdentiPy engine \cite{Levitsky2018} are progressively incorporated to the tool \footnote{\href{https://github.com/compomics/peptide-shaker/issues/309}{https://github.com/compomics/peptide-shaker/issues/309}}. Their usage is predicted to improve identification rates and quality of the matches.

@@

Sample fractionation allows for a deeper analysis of the peptide mix, which in turn leads to increased identifications and an increased amount of data to be analyzed. However, it simultaneously provides an extra random effect that difficults quantification if not handled properly.

### 2018-08-01 thp1.tex
Backup

After preprocessing and missing-data handling, a dataset of 1641 peptides was input to BayesQuant, and 269 proteins were quantified. A histogram of the mean of the inferred posterior distributions manifests the centrality of its distribution, with most values within the predefined \ac{ROPE} of [-0.4, 0.4]. In line with what was observed in the results above, the distribution was found to be skewed toward negative values (see figure \ref{fig:compomics_bq}A). Only four proteins were assigned a 95\% \ac{HPDI} not touching the \ac{ROPE} (see figure \ref{fig:compomics_bq}B and table \ref{tab:thp1_bq_results}). The distribution of the \ac{HPDI} width summarises the overall uncertainty in the estimation process. Most proteins had a very narrow HPDI, as shown in figure \ref{fig:compomics_bq}C. As expected, the width of the interval exhibited some degree of correlation with the mean of the posterior, as shown in figure \ref{fig:compomics_bq}D.

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Sat Jul 21 18:02:39 2018
\begin{table}[!h]
\centering
\begin{tabular}{rlrrrr}
  \hline
 & Protein & log2FC & HPDI start & HPDI end & Peptides \\ 
  \hline
& O75475 & -0.61 & -0.80 & -0.43 & 2 \\ 
   & O15144 & -0.75 & -0.98 & -0.51 & 3 \\ 
   & P02792 & 0.50 & 0.41 & 0.58 & 4 \\ 
   & P04406 & -1.09 & -1.45 & -0.77 & 4 \\ 
   \hline
\end{tabular}
\caption{\ac{DAP} declared by the Compomics+BayesQuant pipeline.}
\label{tab:thp1_bq_results}
\end{table}

@@

On the one hand, nine terms (1, 2, 3, 4, 6, 8, 9, 10 and 17) were cell compartment terms associated with cellular secretion and the extracelullar space. On the other hand, five terms (5, 7, 14, 15, 19) were related to the immunological response. The remaining terms were connected to RNA functions and the nervous system.

@@

\section{Conclusion}
A fine biological interpretation of the results depends on the original question that motivated the study, and thus the work on this chapter did not focus on that, but instead showed how it can be started with the outputsability of the tools presented in the previous chapters. The results indicated that these tools can be deployed to perform computational analyses and biological inference on protein samples with open-source software with no cost for \ac{NZ}. However, the absence of proper missing data handling developed into the inability to quantify hundreds of proteins, signaling what are the next steps to improve the tools.

>>>>

After preprocessing and missing-data handling, a dataset of 1641 peptides was passed to BayesQuant, and 269 proteins were quantified. A histogram of the mean of the inferred posterior distributions manifests its centrality, with most values within the predefined \ac{ROPE} of [-0.4, 0.4]. In line with what was observed in the results above, the distribution was found to be skewed toward negative values (see figure \ref{fig:compomics_bq}A). Only four proteins were assigned a 95\% \ac{HPDI} not touching the \ac{ROPE} (see figure \ref{fig:compomics_bq}B and table \ref{tab:thp1_bq_results}). The distribution of the \ac{HPDI} width summarises the overall uncertainty in the estimation process. Most proteins had a very narrow HPDI, as shown in figure \ref{fig:compomics_bq}C. As expected, the width of the interval exhibited some degree of correlation with the mean of the posterior, as shown in figure \ref{fig:compomics_bq}D.

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Sat Jul 21 18:02:39 2018
\begin{table}[!h]
\centering
\begin{tabular}{rlrrrr}
  \hline
 & Protein & log2FC & HPDI start & HPDI end & Peptides \\ 
  \hline
& O75475 & -0.61 & -0.80 & -0.43 & 2 \\ 
   & O15144 & -0.75 & -0.98 & -0.51 & 3 \\ 
   & P02792 & 0.50 & 0.41 & 0.58 & 4 \\ 
   & P04406 & -1.09 & -1.45 & -0.77 & 4 \\ 
   \hline
\end{tabular}
\caption{\ac{DAP}s declared by the Compomics+BayesQuant pipeline.}
\label{tab:thp1_bq_results}
\end{table}

@@

On the one hand, nine terms (1, 2, 3, 4, 6, 8, 9, 10 and 17) were cell compartment terms associated with cellular secretion and the extracelullar space. On the other hand, five terms (5, 7, 14, 15, 19) were related to the immunological response. The remaining terms were connected to RNA functions.

These results can be better visualized by making use of a treemap (see figure \ref{fig:treemap}), which makes the interpretation of table \ref{tab:gsea} more straightforward.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{revigo_treemap.png}
\caption[REVIGO treemap]{\textbf{REVIGO treemap}. A top-level representation of the enriched GO-terms. Each top-level term is mapped to an area whose size is proportional to the significance of the term. This way, a big area indicates more enrichment \cite{Supek2011}.}
\label{fig:treemap}
\end{figure}

@@

\section{Conclusion}
A fine biological interpretation of the results depends on the original question that motivated the study, and thus this chapter did not focus on that, but instead showed how it can be started with the output of the tools presented in the previous chapters. The results indicated the tools can be deployed to perform computational analyses and biological inference on protein samples with open-source software with no cost for \ac{NZ}. However, the absence of proper missing data handling developed into the inability to quantify hundreds of proteins, signaling what are the next steps to improve the tools.
